---
title:  "This Dashboard Has a Map!"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: scroll
---

```{r, warning=F, message=F}
library(plotly)
library(tidyverse)
library(flexdashboard)
library(DT)
library(crosstalk)
library(sf)
library(leaflet)
library(stringr)
```

```{r}

# Read in the data set that August created
mt_data_wide<-read.csv("mountain_true_lat_long.csv", check.names = FALSE)
#add a unique id column:
mt_data_wide$ID<-c( 
  "x1", "x2", "x3", "x4", "x5", "x6", "x7", "x8", "x9", "x10", 
  "x11", "x12", "x13", "x14", "x15", "x16", "x17", "x18", "x19", "x20", 
  "x21", "x22", "x23", "x24", "x25", "x26", "x27", "x28", "x29", "x30", 
  "x31", "x32", "x33", "x34", "x35", "x36", "x37", "x38", "x39", "x40", 
  "x41", "x42", "x43", "x44", "x45", "x46", "x47", "x48"
)
mt_data_wide<-mt_data_wide%>%relocate(ID, .after=`Lat/Long`)
mt_data_wide<-mt_data_wide[,-c(1,2)]
mt_data_wide<-mt_data_wide%>%relocate(c("Latitude", "Longitude"), .after=`Lat/Long`)
mt_data_wide$Site<-as.factor(mt_data_wide$Site)
```

```{r} 
#first get everything data column encoded as numeric. 
mt_data_wide <- mt_data_wide %>% 
  mutate(across(6:91, as.numeric))
mt_data_wide <- mt_data_wide %>% mutate(across(where(is.numeric), ~replace_na(.x, 0)))
```



```{r}
#we also need a long data set for ggplot:

mt_data_long<-mt_data_wide%>%pivot_longer(
  cols=6:91,
  names_to="Date",
  values_to="Ecoli")
```

```{r, warning=F, message=F}

#I want to create plots and widgets that "talk" to each other.  We need to create "shared" data sets as follows:
mt_sd_wide<-SharedData$new(mt_data_wide,  key=~ID, group="my_dashboard")  
mt_sd_long<-SharedData$new(mt_data_long, key=~ID, group="my_dashboard")

```

```{r}
# WOOOO time series
time_s <- plot_ly(data = mt_sd_long, x = ~Date, y = ~Ecoli,
               type = 'scatter', mode = 'lines', color=~Site) %>%
  layout(title = "Time Series",
         xaxis = list(title = "Date"),
         yaxis = list(title = "E. Coli"),
         showlegend = FALSE
         #add_segments(x = ~Date, xend = ~Date,
           #y = 135, yend = 135, 
           #line = list(color = "red", dash = "dash"),
           #showlegend = FALSE) 
)
```


```{r}
map<-leaflet(mt_sd_wide) %>%
  addTiles() %>%
  addMarkers(
    lng = ~Longitude, 
    lat = ~Latitude,
    label=~paste0(Site)
  )
  
```

 



## {.sidebar}

Here is a basic   **dashboard**

  + Goal:  Visualize ecoli data
  + Status:  Not  AT ALL complete
  
```{r}
filter_select("site", "Select Site:", mt_sd_long, ~Site)

```
  

## Column

### This is a map -- when fully up to date this should be color coded based on water quality, hopefully clickable? i think shiny is the way to go with this because i really dont like the crosstalk replacements for it

```{r}
map
```
## Column

### These are boxplots: 

here should be time series data, with the actual ecoli v predictions + confidence interval for each site -- able to be filtered by site
```{r}
time_s
```



### This is a datatable

here should be the most recent e coli measurement, our prediction, what category it falls in, the formula we used (and maybe what each of the predictor variable values are?)
