---
title:  "Warren Wilson EColi Regression!!!!"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    runtime: shiny
    theme:
      bg: "#FDFDFD"
      fg: "#101712"
      navbar-bg: "#98b39f"
---
```{r, warning=F, message=F}
library(crosstalk)
library(DT)
library(flexdashboard)
library(leaflet)
library(plotly)
library(sf)
library(shiny)
library(stringr)
library(tidyverse)

source("functions.R")
```

```{r}
# data from predicting.Rmd
mt_data_wide<-read.csv("data/dashboard_data.csv", check.names = FALSE)

mt_data_wide<-mt_data_wide %>% mutate(ID = paste0("x", cur_group_id()), .by = Site) %>% relocate(ID, .after=Site)
mt_data_wide$Site<-as.factor(mt_data_wide$Site)

mt_data_wide <- mt_data_wide %>% mutate(across(6:last_col(), ~ round(., 2)))
```

```{r}
#we also need a long data set for ggplot: 
mt_data_long<-mt_data_wide %>% pivot_longer(cols=6:last_col(), names_to="Date", values_to="Ecoli") %>%
  mutate(Date = as.Date(Date))
```

```{r}
# very roundabout way to get column names since they will change everyday -- 
# these column names are for the textbox mainly  
today_col_wide <- last(colnames(mt_data_wide))

mt_data_wide <- add_groups(mt_data_wide, today_col_wide)
today_group_col_wide <- last(colnames(mt_data_wide))

mt_data_long <- add_groups(mt_data_long, "Ecoli")
mt_data_long$Ecoli.GROUP <- as.factor(mt_data_long$Ecoli.GROUP)
```

```{r}
# reactive variables and filtered data and such
selected_site <- reactiveVal(NULL)
marker_just_clicked <- reactiveVal(FALSE)

observeEvent(input$mymap_marker_click, {
  req(input$mymap_marker_click$id)
  marker_just_clicked(TRUE)
  selected_site(input$mymap_marker_click$id)
})

observeEvent(input$mymap_click, {
  if (marker_just_clicked()) {
    marker_just_clicked(FALSE)
    return()
  }
  selected_site(NULL)
})

filtered_long <- reactive({
  site_id <- selected_site()
  if (is.null(site_id)) {
    return(mt_data_long) 
  } else {
    return(mt_data_long %>% filter(ID == site_id))
  }
})

filtered_wide <- reactive({
  site_id <- selected_site()
  if (is.null(site_id)) {
    return(mt_data_wide)
  } else {
    return(mt_data_wide %>% filter(ID == site_id))
  }
})
```

```{r}
output$mymap <- renderLeaflet({
  pal <- colorFactor(palette = c("primary" = "limegreen", "secondary" = "#f1c40f", "unsafe" = "orangered"), 
                     domain = mt_data_wide[[today_group_col_wide]])
  
  leaflet(mt_data_wide) %>% 
    addTiles() %>%
    addProviderTiles(providers$Stadia.OSMBright) %>%
    addCircleMarkers(
      lng = ~Longitude, 
      lat = ~Latitude,
      label = ~Site,
      layerId = ~ID,
      fillOpacity = 1, 
      color = ~pal(mt_data_wide[[today_group_col_wide]]))
})
```

```{r}
# i definitely did let ai figure out the buttons but they do work
output$site_buttons <- renderUI({
  sites <- mt_data_wide %>% select(ID, Site) %>% distinct()
  
  button_list <- lapply(seq_len(nrow(sites)), function(i) {
    site_id <- sites$ID[i]
    site_name <- sites$Site[i]
    
    actionButton(
      inputId = paste0("btn_", site_id), 
      label = site_name,
      style = "color: #fff; background-color: #98b39f; border-color: #00000000; font-size: 16px; outline: none;
      border-radius: 5px")
  })
  
  div(style = "display: flex; flex-wrap: wrap; gap: 10px;", button_list)
})

observe({
  sites <- mt_data_wide %>% select(ID) %>% distinct() %>% pull(ID)
  
  for (site_id in sites) {
    local({
      id <- site_id
      btn_id <- paste0("btn_", id)
      
      observeEvent(input[[btn_id]], {selected_site(id)})
    })
  }
})
```

```{r}
output$timeseries <- renderPlotly({
  df = filtered_long() %>% na.omit()
  mycolors = c("primary" = "limegreen", "secondary" = "#f1c40f", "unsafe" = "orangered")
  marker_colors <- unname(mycolors[as.character(df$Ecoli.GROUP)])
  
  plot_ly(data = df) %>% 
    add_lines(data = df %>% filter(Variable == "E.Coli.SUM"), 
              x = ~Date, y = ~Ecoli, name = "Sampled E Coli",
              line = list(color = "darkgreen")) %>%
    add_lines(data = df %>% filter(Variable == "Predict"), 
              x = ~Date, y = ~Ecoli, name = "Predicted E Coli",
              line = list(color = "orangered")) %>%
    add_markers(data = df, x = ~Date, y = ~Ecoli, marker = list(size = 8, color = marker_colors),
                text = ~paste(Variable), hoverinfo = "text+x+y", legendgroup = ~Ecoli.GROUP, 
                inherit = FALSE, name = "Safety Category") %>%
    layout(title = "Time Series",
           xaxis = list(
             title = "Date",
             range = c("2026-05-01", "2026-08-01")),
           yaxis = list(title = "E. Coli", range = c(0, 5000)),
           showlegend = TRUE)
})
```

```{r}
output$site_summary <- renderText({
  predict_row <- filtered_wide() %>% filter(Variable == "Predict")
  actual_row <- filtered_wide() %>% filter(Variable == "E.Coli.SUM") %>% select(-last_col()) 
  # last column being groups of the predicted values, which we do not need for the actual row
  
  if (nrow(predict_row) > 2) {
    return("These predictions are made through logarithmic linear regression models. Each site was tested with an accuracy result of between 70-85%. Swim at your own risk!")
    }
  
  site_name <- predict_row$Site
  today_prediction <- predict_row[[today_col_wide]]
  today_group <- predict_row[[today_group_col_wide]]
  last_real_value <- actual_row %>% select(where(~ !any(is.na(.)))) %>% select(last_col())
  last_real_date <- names(last_real_value)
  
  my_text <- paste0("Predicted E. coli: ", today_prediction, " MPN/100mL", "\n",
                    "Today's safety category is ", today_group, ".", "\n", "Last recorded value: ",
                    last_real_value, " CFU/100mL, recorded on ", last_real_date, ".")
  
  if(is.na(today_prediction)) {
    return(paste0("Unfortunary, we do not have a prediction for today :( ", "\n", "The last recorded value was ",
           last_real_value, " CFU/100mL, recorded on ", last_real_date, "."))
  }
  
  return(my_text) 
})

output$site_header <- renderUI({
  predict_row <- filtered_wide() %>% filter(Variable == "Predict")
  
  site_name <- predict_row$Site
  today_group <- predict_row[[today_group_col_wide]][1]
  
  colors <- c("primary" = "#2ecc71", "secondary" = "#f1c40f", "unsafe" = "#e74c3c")
  
  bg_color <- colors[today_group]
  
  if (length(bg_color) == 0){
    bg_color <- "#00000000"
  }
  
  if (nrow(predict_row) > 2) {
    return("Please click on a site to view more information!")
    }
  
  return(
    div(style = paste0("background-color: ", bg_color, "; color: white; padding: 10px;"), site_name))
})
```
 



## {.sidebar data-width=300}

```{r}
uiOutput("site_buttons")
hr()

observeEvent(input$clear_selection, {
  selected_site(NULL)
})
```

Primary recreation is defined as E. Coli levels less than 126 MPN/100mL, meaning it is recommended safe to swim and be submerged in the water.

Secondary recreation is defined as E. Coli levels between 126 and 886 MPN/100mL, where the water is not recommended for swimming but generally safe for boating or paddling, with a lower potential for ingestion of the water. 

Above 886 MPN/100mL is classified as unsafe, and it is not recommended to be in the water at this time. 

## Column

### This is a map

```{r}
leafletOutput("mymap")
```

## Column 

### {data-height=225}

```{r}
h2(uiOutput("site_header"))

# textOutput("site_summary") needs to be wrapped in CSS to read the line breaks for some odd reason
div(style = "white-space: pre-wrap;", textOutput("site_summary"))
```

### My time series!! 
```{r}
plotlyOutput("timeseries")
```
