---
title: "predicting!!"
author: "August Duncan"
date: "2026-07-14"
output: html_document
---

```{r}
library(tidyverse)
source("functions.R")
```

okay this is going to predict the ecoli values for the past to compare with the future, as well as predicting today. 

there four sites at the moment, but this should be fairly easy to just add more if and when we get there. 

formulas to be applied (these have been previously worked on and generated):

Pearson Bridge -- E.Coli.LOG ~ -3.05 + 0.11* Turbidity.MEAN.ROOT + 0.91* Discharge.MEAN.LOG - 0.22* Dissolved.oxygen.MAXIMUM + 0.25* Precip.2.SUM + 0.56*pH.MINIMUM
  USGS Site: USGS-03451500, Mountain True site "FB at Pearson Bridge"
    
Hominy Creek Greenway -- E.Coli.LOG ~ 3.61 + 0.57* Turbidity.MEAN.LOG + 0.26* Precip.2.SUM +  0.34* Gage.height.MEAN - 0.26*Dissolved.oxygen.MAXIMUM

Bacoate -- E.Coli.LOG ~ 6.86 + 0.002 * Turbidity.RANGE + 0.85 * Precip.1.SUM - 0.06 * Specific.cond.at.25C.MINIMUM - 0.24 * Dissolved.oxygen.MEAN

Westfeldt Park -- E.Coli.LOG ~ 0.31 + 0.05* Turbidity..Form.Neph.MINIMUM + 0.64* Discharge.MEAN.LOG - 0.06* Precip.1.SUM - 0.09*Precip.7.SUM


this gives us all the actual data from mountain true, what we would have predicted for those days, and today's predicted value is the last column. 

```{r}
today <- as.character(Sys.Date() -1 )
```

```{r, warning=FALSE, message=FALSE}
#pearson bridge
pb <- get_the_data_and_precip("USGS-03451500", "2020-01-01", today, "FB at Pearson Bridge")
pb_today <- get_todays_data("USGS-03451500")
pb <- bind_rows(pb, pb_today)

pb <- pb %>% mutate(across(c(where(is.numeric), -E.Coli.SUM), ~ replace_na(., median(., na.rm = TRUE))))


pb$Predict <- 10^(-3.05 + 0.11 * sqrt(pb$Turbidity..Form.Neph.MEAN) + 0.91* log10(pb$Discharge.MEAN) - 0.22* pb$Dissolved.oxygen.MAXIMUM + 0.25* pb$Precip.2.SUM + 0.56*pb$pH.MINIMUM)

pb <- pb %>% select(Date, E.Coli.SUM, Predict) %>% pivot_longer(cols = c(E.Coli.SUM, Predict), names_to = "Variable", values_to = "Value") %>% pivot_wider(names_from = Date, values_from = Value) %>% mutate(Site = "FB at Pearson Bridge")
```

```{r, warning=FALSE, message=FALSE}
# hominy creek
hominy <- get_the_data("USGS-0344878100", "2020-01-01", today, "Hominy Creek Greenway")
hominy_today <- get_todays_data("USGS-0344878100")
hominy <- bind_rows(hominy, hominy_today)

#noaa precip does not get updated frequently enough, yet usgs precipitation only has data since 2024....
hominy <- add_noaa_precip(hominy, "2020-01-01", today)

hominy <- hominy %>% mutate(across(c(where(is.numeric), -E.Coli.SUM), ~ replace_na(., median(., na.rm = TRUE))))


hominy$Predict <- 10^(3.61 + 0.57*log10(hominy$Turbidity..Form.Neph.MEAN) + 0.26*hominy$Precip.2.SUM +  0.34* hominy$Gage.height.MEAN - 0.26*hominy$Dissolved.oxygen.MAXIMUM)

hominy <- hominy %>% select(Date, E.Coli.SUM, Predict) %>% pivot_longer(cols = c(E.Coli.SUM, Predict), names_to = "Variable", values_to = "Value") %>% pivot_wider(names_from = Date, values_from = Value) %>% mutate(Site = "Hominy Creek Greenway")
```


```{r, warning=FALSE, message=FALSE}
# bacoate
bacoate <- get_the_data("USGS-03451200", "2020-01-01", today, "Bacoate")
bacoate_today <- get_todays_data("USGS-03451200")
bacoate <- bind_rows(bacoate, bacoate_today)

bacoate <- add_noaa_precip(bacoate, "2020-01-01", today)
bacoate <- bacoate %>% mutate(across(c(where(is.numeric), -E.Coli.SUM), ~ replace_na(., median(., na.rm = TRUE))))


bacoate$Predict <- 10^(6.86 + 0.002*(bacoate$Turbidity..Form.Neph.MAXIMUM - bacoate$Turbidity..Form.Neph.MINIMUM) + 0.85*bacoate$Precip.1.SUM - 0.06*bacoate$Specific.cond.at.25C.MINIMUM - 0.24 * bacoate$Dissolved.oxygen.MEAN)

bacoate <- bacoate %>% select(Date, E.Coli.SUM, Predict) %>% pivot_longer(cols = c(E.Coli.SUM, Predict), names_to = "Variable", values_to = "Value") %>% pivot_wider(names_from = Date, values_from = Value) %>% mutate(Site = "Bacoate")
```

```{r, warning=FALSE, message=FALSE}
# westfeldt park
westfeldt <- get_the_data("USGS-03447687", "2020-01-01", today, "FB at Westfeldt Park")
westfeldt_today <- get_todays_data("USGS-03447687")
westfeldt <- bind_rows(westfeldt, westfeldt_today)

westfeldt <- add_usgs_precip(westfeldt, "2020-01-01", today, "USGS-352536082333245")
westfeldt <- westfeldt %>% mutate(across(c(where(is.numeric), -E.Coli.SUM), ~ replace_na(., median(., na.rm = TRUE))))


westfeldt$Predict <- 10^(0.31 + 0.05*westfeldt$Turbidity..Form.Neph.MINIMUM + 0.64*log10(westfeldt$Discharge.MEAN) - 0.06*westfeldt$Precip.1.SUM - 0.09*westfeldt$Precip.7.SUM)

westfeldt <- westfeldt %>% select(Date, E.Coli.SUM, Predict) %>% pivot_longer(cols = c(E.Coli.SUM, Predict), names_to = "Variable", values_to = "Value") %>% pivot_wider(names_from = Date, values_from = Value) %>% mutate(Site = "FB at Westfeldt Park")
```

```{r}
combined_data <- bind_rows(pb, hominy, bacoate, westfeldt)
combined_data <- combined_data %>% select(order(names(.))) %>% relocate(Site, Variable) 
```

so theres definitely some issues on access to todays data (usgs hates me) but i have todays prediction for pearson bridge at least! 

```{r}
combined_data <- add_lat_long(combined_data)
```

i dont know how this data should be formatted in order to be parsed to the dashboard. should i have one dataframe
being actual values and one being predicted? 


```{r}
write.csv(combined_data, "dashboard_data.csv", row.names = FALSE)
```








