---
title: "predicting!!"
author: "August Duncan"
date: "2026-07-14"
output: html_document
---

```{r}
library(tidyverse)
source("functions.R")
```

okay this is going to predict the ecoli values for the past to compare with the future, as well as predicting today. 

we have four sites at the moment, but this should be fairly easy to just add more if and when we get there. 

formulas to be applied (these have been previously worked on and generated):

Pearson Bridge -- E.Coli.LOG ~ -3.05 + 0.11* Turbidity.MEAN.ROOT + 0.91* Discharge.MEAN.LOG - 0.22* Dissolved.oxygen.MAXIMUM + 0.25* Precip.2.SUM + 0.56*pH.MINIMUM
  USGS Site: USGS-03451500, Mountain True site "FB at Pearson Bridge"
    
Hominy Creek Greenway -- E.Coli.LOG ~ 3.61 + 0.57* Turbidity.MEAN.LOG + 0.26* Precip.2.SUM +  0.34* Gage.height.MEAN - 0.26*Dissolved.oxygen.MAXIMUM

Bacoate -- E.Coli.LOG ~ 6.86 + 0.002 * Turbidity.RANGE + 0.85 * Precip.1.SUM - 0.06 * Specific.cond.at.25C.MINIMUM - 0.24 * Dissolved.oxygen.MEAN

Westfeldt Park -- E.Coli.LOG ~ 0.31 + 0.05* Turbidity..Form.Neph.MINIMUM + 0.64* Discharge.MEAN.LOG - 0.06* Precip.1.SUM - 0.09*Precip.7.SUM


this gives us all the actual data from mountain true, what we would have predicted for those days, and today's predicted value is the last column. 
```{r, warning=FALSE, message=FALSE}
# this will have the updated 2026 mountain true data when i get around to adding that in. 

#pearson bridge
pb <- get_the_data_and_precip("USGS-03451500", "2020-01-01", "2025-12-31", "FB at Pearson Bridge")
pb_today <- get_todays_data("USGS-03451500")

pb <- pb %>% mutate(across(where(is.numeric), ~ replace_na(., median(., na.rm = TRUE))))
pb <- bind_rows(pb, pb_today)

pb$Predict <- 10^(-3.05 + 0.11 * sqrt(pb$Turbidity..Form.Neph.MEAN) + 0.91* log10(pb$Discharge.MEAN) - 0.22* pb$Dissolved.oxygen.MAXIMUM + 0.25* pb$Precip.2.SUM + 0.56*pb$pH.MINIMUM)

pb <- pb %>% select(Date, E.Coli.SUM, Predict) %>% pivot_longer(cols = c(E.Coli.SUM, Predict), names_to = "Variable", values_to = "Value") %>% pivot_wider(names_from = Date, values_from = Value)
```

```{r, warning=FALSE, message=FALSE}
# hominy creek
hominy <- get_the_data("USGS-0344878100", "2020-01-01", "2025-12-31", "Hominy Creek Greenway")
hominy_today <- get_todays_data("USGS-0344878100")
hominy <- bind_rows(hominy, hominy_today)

hominy <- add_noaa_precip(hominy, "2020-01-01", as.character(Sys.Date()))

hominy <- hominy %>% mutate(across(c(where(is.numeric), -E.Coli.SUM), ~ replace_na(., median(., na.rm = TRUE))))


hominy$Predict <- 10^(3.61 + 0.57*log10(hominy$Turbidity..Form.Neph.MEAN) + 0.26*hominy$Precip.2.SUM +  0.34* hominy$Gage.height.MEAN - 0.26*hominy$Dissolved.oxygen.MAXIMUM)

hominy <- hominy %>% select(Date, E.Coli.SUM, Predict) %>% pivot_longer(cols = c(E.Coli.SUM, Predict), names_to = "Variable", values_to = "Value") %>% pivot_wider(names_from = Date, values_from = Value)
```



```{r, warning=FALSE, message=FALSE}
# bacoate
bacoate <- get_the_data("USGS-0344878100", "2020-01-01", "2025-12-31", "Hominy Creek Greenway")
bacoate_today <- get_todays_data("USGS-0344878100")
bacoate <- bind_rows(bacoate, bacoate_today)

bacoate <- add_noaa_precip(bacoate, "2020-01-01", "2025-12-31")
bacoate <- bacoate %>% mutate(across(where(is.numeric), ~ replace_na(., median(., na.rm = TRUE))))


hominy$Predict <- 10^(3.61 + 0.57*log10(hominy$Turbidity..Form.Neph.MEAN) + 0.26*hominy$Precip.2.SUM +  0.34* hominy$Gage.height.MEAN - 0.26*hominy$Dissolved.oxygen.MAXIMUM)

hominy <- hominy %>% select(Date, E.Coli.SUM, Predict) %>% pivot_longer(cols = c(E.Coli.SUM, Predict), names_to = "Variable", values_to = "Value") %>% pivot_wider(names_from = Date, values_from = Value)
```




