---
title: "predicting!!"
author: "August Duncan"
date: "2026-07-14"
output: html_document
---

```{r}
library(tidyverse)
source("functions.R")
```

okay this is going to predict the ecoli values for the past to compare with the future, as well as predicting today. 

we have four sites at the moment, but this should be fairly easy to just add more if and when we get there. 

formulas to be applied (these have been previously worked on and generated):

Pearson Bridge -- E.Coli.LOG ~ -3.05 + 0.11* Turbidity.MEAN.ROOT + 0.91* Discharge.MEAN.LOG - 0.22* Dissolved.oxygen.MAXIMUM + 0.25* Precip.2.SUM + 0.56*pH.MINIMUM
  USGS Site: USGS-03451500, Mountain True site "FB at Pearson Bridge"
    
Hominy Creek Greenway -- E.Coli.LOG ~ 3.61 + 0.57* Turbidity.MEAN.LOG + 0.26* Precip.2.SUM +  0.34* Gage.height.MEAN - 0.26*Dissolved.oxygen.MAXIMUM

Bacoate -- E.Coli.LOG ~ 6.86 + 0.002 * Turbidity.RANGE + 0.85 * Precip.1.SUM - 0.06 * Specific.cond.at.25C.MINIMUM - 0.24 * Dissolved.oxygen.MEAN

Westfeldt Park -- E.Coli.LOG ~ 0.31 + 0.05* Turbidity..Form.Neph.MINIMUM + 0.64* Discharge.MEAN.LOG - 0.06* Precip.1.SUM - 0.09*Precip.7.SUM

```{r}
pb <- get_the_data_and_precip("USGS-03451500", "2020-01-01", "2025-12-31", "FB at Pearson Bridge")

pb <- pb %>% mutate(across(where(is.numeric), ~ replace_na(., median(., na.rm = TRUE))))

pb$Predict <- 10^(-3.05 + 0.11 * sqrt(pb$Turbidity..Form.Neph.MEAN) + 0.91* log10(pb$Discharge.MEAN) - 0.22* pb$Dissolved.oxygen.MAXIMUM + 0.25* pb$Precip.2.SUM + 0.56*pb$pH.MINIMUM)

pb <- pb %>% select(Date, E.Coli.SUM, Predict) %>% pivot_longer(cols = c(E.Coli.SUM, Predict), names_to = "Variable", values_to = "Value") %>% pivot_wider(names_from = Date, values_from = Value)
```





